Reflections on reflection for \\ intuitionistic set theories

### 2019-01-11 paper-reflection.tex
Baby form of the reflection theorem

\begin{rem}In the special case~$n = 1$, the proof of Lemma~\ref{lemma:izf-smallstep} can
be simplified to only use~$\RRS$ instead of~$\RRS_2$.\end{rem}

@@

% XXX: State mini-reflection theorem.

@@

\textbf{XXX: Use mini-reflection theorem instead.}
By Theorem~\ref{thm:izf-bigstep}, reflection is available; hence there is a
set~$S \supseteq \{x_0\}$ which reflects the three formulas~``$x \in X$'', ``$\langle x,y
\rangle \in R$'' and~``$\forall x\_ x \in X \Rightarrow \exists y\_ (y \in X \wedge \langle x,y
\rangle \in R)$''.

>>>>

\begin{rem}\label{rem:izf-smallstep-1}
In the special case~$n = 1$, the proof of Lemma~\ref{lemma:izf-smallstep} can
be simplified to only use~$\RRS$ instead of~$\RRS_2$, because in this case it
suffices for the set~$B$ to be such that for any~$H \in B$, there is a set~$H'
\in B$ such that~$H$ and~$H'$ are related as in the conclusion of
Lemma~\ref{lemma:izf-microstep}.
\end{rem}

@@

\begin{scholium}\label{scholium:izf-bigstep-1}
Let~$\varphi(x_1,\ldots,x_n)$ be a formula in the language of set
theory with free variables as indicated. Assume that the surrounding scope of
any unbounded quantifier in~$\varphi$ contains at most one free variable.
Then~$\IZF+\RRS$ proves the same conclusion as stated in
Theorem~\ref{thm:izf-bigstep}.
\end{scholium}

\begin{proof}The condition on the number of free variables allows the proof of
Theorem~\ref{thm:izf-bigstep} to be adapted to employ the version of
Lemma~\ref{lemma:izf-smallstep} outlined in Remark~\ref{rem:izf-smallstep-1},
which requires only~$\IZF+\RRS$ instead of~$\IZF+\RRS_2$.
\end{proof}

@@

We cannot apply Scholium~\ref{scholium:izf-bigstep-1} to the formula~``$\forall
x \in X\_ \exists y \in X\_ \langle x,y \rangle \in R$'' since unbounded
quantifiers implicitly appearing in the formulas~``$x \in X$'' and~``$\langle
x,y \rangle \in R$'' (recalling that~$X$ and~$R$ are classes) may violate the
condition on the number of free variables. However, we can opt to leave these
subformulas to be untranslated when carrying out the~$S$-relativization; with
this understanding, Scholium~\ref{scholium:izf-bigstep-1} can be applied to
yield a set~$S \ni x_0$ such that
\[ \forall x \in S\_ (x \in X \Rightarrow \exists y \in S\_
  (y \in X \wedge \langle x,y \rangle \in R)). \]
